### 2018-05-24 paper.tex
adding details to JSExplain, WIP

%\documentclass{article}
\documentclass[twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{positioning,fit,shapes,arrows.meta}

\author{ K\'evin Le Bon \and Alan Schmitt }

\title{MLExplain}

\begin{document}
\maketitle

\begin{abstract}
  MLExplain is a step-by-step interpreter for OCaml that enables the
  user to inspect both their program's state and the interpreter's state
  itself. This interpreter aims to provide the user with a better understanding
  of the semantics of OCaml semantics.
\end{abstract}

\section{Introduction}

The semantics of a programming language can be very complex. When a language
has no specification, the semantics is then defined by the implementations of
the language, i.e., interpreters and compilers. However even when a
specification is available, it can be difficult to understand why the execution
of a specific program results to a certain output.

The goal of the \emph{JSExplain} project \cite{chargueraud:hal-01745792} is to
describe the execution of a JavaScript program by showing every step of an
interpreter whose behavior is very closest to the specification of JavaScript.
In this paper, we show how we adapted JSExplain to OCaml.

Unlike JavaScript, OCaml has no official specification. However, OCaml code
execution is fairly simple because the amount of language constructions is
small. We have written an interpreter for OCaml's typed abstract syntax tree
(AST). This AST is close to the source code and need little transformations but
supplies us with resolved names -- which is useful for module and signatures
inclusion and absolutely mandatory for features like named parameters and
optional parameters in functions. The semantic we give to OCaml is higher level
than the one described in \emph{ZINC} \cite{Leroy-ZINC} virtual machine which is
at the same level as bytecode.

\section{JSExplain}

\subsection{Double Debugger}
\label{subsec:double-debugger}

We call a \emph{double debugger} an interpreter that is able to run step by step
and display the program's state and current location in source, as well as the
state and location in the source of the interpreter itself. JSExplain is a
double debugger for JavaScript. The aim of this tool is to help a user to better
understand JavaScript's semantics, by providing an interpreter that is very
close to the specification of JavaScript. JSExplain's interface highlights the
expression of the program currently evaluated as well as the instruction of the
interpreter that is executed.

\subsection{Architecture}

The core of JSExplain is a JavaScript interpreter written in a subset of OCaml.
This interpreter is compiled to a subset of JavaScript and instrumented to
generate a trace of its execution when run. This trace can then be navigated
using a web-based
tool.\footnote{https://jscert.github.io/jsexplain/branch/master/driver.html}

The subset of OCaml we support for writing the interpreter is purely functional
with variables, constants, sequence, conditional, let-binding, function
definition, function application (with support for prefix and infix functions),
data constructors, records (including record projections, and the
``record-with'' construct to build a copy of a record with a number of fields
updated), tuples (i.e., anonymous records), and simple pattern matching (only
with non-nested patterns, restricted to data constructors, constants, variables,
and wildcards). For convenience, let-bindings and functions may bind simple
patterns (as opposed to only variables). We also support \texttt{ppx} extensions
for monadic programming, to simplify the handling of errors.

The target language of our compiler is a purely functional subset of JavaScript
where there is never any type conversion, where objects are used as records (no
use of their \texttt{prototype} field), and where tuples are encoded into
arrays.

ALAN: stopped here

JSExplain is a visual interpreter. To display its state, the interpreter is
compiled with a compiler that adds calls to a tracing function. The traced
retrieved during the execution of the program gives many information about the
state of the interpreter:

\begin{itemize}
  \item existing variables and their values
  \item return values of functions of the interpreter
\end{itemize}

JSExplain's web interface uses these traces in order to display the program and
interpreter state at every step of the execution.

The whole project takes the shape of a web application composed of the
JavaScript interpreter and a view. The view allows the user to write JavaScript
code and run it. It also displays the state of the program and the state of the
interpreter.

The program's execution is not interactive: when the button \texttt{run} is
pressed, the program is executed all at one and the traces are generated. Only
then, the user can replay the execution step by step.

\subsection{JSExplain's compiler}

% Référence pour coq ?
JSExplain was first written in Coq. OCaml code has been extracted from the Coq
source and is now the primary source code. The interpreter is compiled into
JavaScript in order to use it within a web page. For this reason, the project
embeds a homemade OCaml-to-JavaScript compiler.

JSExplain also needs its own OCaml to JavaScript compiler in order to place the
calls to \verb|log_event| which is the tracing function. This function is also
used to get the location of the different syntactical objects in the source
file in order to produce code highlighting in JSExplain's interface.

This compiler offers several compilation modes:

\begin{itemize}
  \item \texttt{unlog}: generate JavaScript without tracing.
  \item \texttt{log}: generate JavaScript with tracing.
  \item \texttt{mlloc}, \texttt{token} and \texttt{ptoken}: generate files with
    informations about syntastical elements from the input file.
  \item \texttt{pseudo}: generate JavaScript enhanced with syntactic sugar for
    pattern matching or monadic operations -- it is supposed to be easier to
    read.
\end{itemize}

Modes \texttt{unlog} and \texttt{log} exist so that auxiliary code in the
interpreter -- like linked-list or hash-map internal-management functions -- is
not traced.  The user is thus able to concentrate on the logical part of the
interpreter functioning and forget about implementation details.

\subsubsection{fOCaml}
This compiler does not recognized the whole OCaml language, it can only handle
a small subset we named \texttt{fOCaml}. Among other restrictions, it cannot
handle objects and imperative features, exceptions, inner modules, and tuples.
The main reason for this limitation is that compiling OCaml into JavaScript is
not trivial, many features -- e.g. nested pattern matching -- are hard to
represent to JavaScript.

The fOCaml code is compiled into \texttt{fJs}, a functional subset of
JavaScript. The compiler preserves as much as possible the code structure to
enable the user to understand both fOCaml and fJs codes as the same one.

% Petite explication sur le fait que du coup on écrit du code fonctionnel pur
% et les implications que ça a ?

\section{MLExplain}

At first glance, JSExplain seems tightly linked to JavaScript. However, the
architecture of the project actually allows us to take off the interpreter
completely and replace it with another one.

% Plus de blabla

\subsection{Two compilers}

To interpret OCaml code, we need to parse it. However, our compiler does not
allow us to use the standard library nor an external one. Fortunately,
JSExplain itself did not trace the parsing process. That is why we have been
able to separate totally the parser from the interpreter and use another
compiler to compile it. The figure \ref{fig:archi} describes the architecture
of the whole application.

The parser is compiled with
\emph{js\_of\_ocaml} \cite{DBLP:journals/spe/VouillonB14}, the compiler of the
project \emph{Ocsigen}.

\begin{figure*}[ht]
  \centering
  \begin{tikzpicture}[nodes = {align = center}]
    \node (focaml) [draw, text width=2cm] {Interpreter and libraries (OCaml)};

    \node (fjs) [draw, right = 1cm of focaml, text width=2cm, rectangle split,
    rectangle split parts=2, text centered] {Libraries (JS) \nodepart{second}
      Interpreter with traces (JS)};

    \node (ast) [draw, above = of fjs, text width=2cm] {AST of interpreted
      program};

    \node (source) [draw, right = 2cm of ast, text width=2cm, text centered]
    {interpreted program};

    \node [draw, dotted, fit=(fjs) (ast) (source)] {web page};

    \node (trace) [draw, right = 2cm of fjs, text width=2cm, text centered]
    {trace};

    \path[->,thick,>=stealth] (focaml.north
    east) edge[bend left] node[above left] {compiler} (fjs.north west) ;

    \path[->,thick,>=stealth] (focaml.south east) edge[bend right] node[below
    left] {\parbox{2cm}{\centering tracing compiler}} (fjs.south west);

    \path[->,thick,dotted,>=stealth] (source) edge node [above] {Parser} node
    [below] {Js\_of\_ocaml} (ast);

    \path[->,thick,dotted,>=stealth] (ast) edge (fjs);

    \path[->,thick,dotted,>=stealth] (ast) edge (fjs);

    \path[->,thick,dotted,>=stealth] (fjs) edge (trace);
  \end{tikzpicture}

  \caption{Architecture of MLExplain}
  \label{fig:archi}
\end{figure*}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}